\documentclass[twoside,a4paper]{master}
\usepackage[kerning,spacing]{microtype}
\usepackage{float}
\microtypecontext{spacing=nonfrench}

% Helpfull command to put reminders in your work
\newcommand{\Todo}[1]%
{\fcolorbox{black}{orange}{\parbox{\dimexpr\textwidth-2\fboxsep-2\fboxrule}{\textbf{Todo: #1}}}}
% it uses one argument which is referred to by #1

\newcommand{\Alphabet}{\mathcal{A}\xspace}

% Add bibliography resource
\addbibresource{references.bib}

\begin{document}

% use roman page numbers before the real document starts.
\pagenumbering{roman}

%Change the values here to get your title page and assertion page right
\AdvisorA{Prof. Dr. Matthias Rarey}
\AdvisorB{Prof. Dr. Andrew Torda}
\author{Marius R\"uve}
\MatrikelNo{7657327}
%\AuthorAdress{Musterstrasse 1}{07770 M\"unchhausen}
\title{
  \begin{tabular}{c}
    \textbf{Ligand Transplantation and Optimization} \\[3mm]
    \textbf{in Protein Models}
  \end{tabular}}
\Maketitle

\addchap*{Abstract}
\small
% High-priority thesis TODOs
\Todo{Fill a 200--250 word abstract summarizing problem, method (FoldFusion), key results (quantitative), limitations, and contributions.}
\Todo{Ensure advisor names/titles and matriculation number are correct and match university records.}
\Todo{Decide on final chapter structure: currently Introduction, Methods, Results, Conclusion. Remove references to non-existent chapters (Background, Discussion) or add those chapters.}
\Todo{Move final figures into thesis/figures as PDF/SVG (preferred) or 300 DPI PNG; reference them in Methods/Results.}
\Todo{Add an Appendix section with exact reproducibility details: environment (Python version, key packages), commands to run pipeline, git commit hash, dataset versions, and config.toml snapshot.}
\Todo{Run full LaTeX compile and resolve all warnings: no undefined refs/citations, consistent figure/table numbering, proper hyphenation.}
% TODO: Replace placeholder with final abstract
% Suggested structure: 1) Context & gap, 2) FoldFusion approach, 3) Core results (n targets, transplant success metrics), 4) Implications/limitations.
Test text

\normalsize

\setcounter{tocdepth}{1}
\tableofcontents

\pagenumbering{arabic}
\include{chapters/01_introduction}
\include{chapters/02_methods}
\include{chapters/03_results}
\include{chapters/04_conclusion}

\newpage
\small
% Bibliography using biblatex
\printbibliography[title=References]
\normalsize

% appendix

\appendix
\noappendicestocpagenum
\addappheadtotoc

\Chapter{Reproducibility Details}
\Todo{Insert environment details: OS, Python version, key packages with versions, hardware.}
\Todo{Insert exact commands (Makefile targets or CLI) used to generate results and figures.}
\Todo{Record git commit hash and config snapshot used for final evaluation.}

\Chapter{Supplementary Tables and Figures}
\Todo{Add tables summarizing dataset proteins, donors, ligands, and transplant success by category.}
\Todo{Include any extended figures not in the main text.}


\Assertion
\end{document}
